A Sound and Complete Deep Embedding of Arbitrary Normal Modal Logics in Coq

### 2025-08-20 paper/COLA/paper.tex
Justify Löb's theorem

In this section we describe a case study of the defined modal logic library:
a proof L\"ob theorem in an extension of system K4 with modal fixed points.
The importance of L\"ob's theorem is due to its use in provability
logic~\cite{boolos_1994}, used to abstract details from encodings in G\"odel's
incompleteness results.

In order to state L\"ob's theorem we need a way to encode modal
fixed points. Formally, let $\varphi(X)$ be a formula with only one propositional variable
$X$. The modal fixed point of $\varphi$ is then a formula $\Psi$ such that
$\Psi \leftrightarrow \varphi(\Box \Psi)$.
We represent the existence of modal fixed points for all one-variable formulas by a definition \mintinline{coq}{fixed_point},
which is parameterized by a system of modal logic, \mintinline{coq}{A}, and a
theory, \mintinline{coq}{G}, and encodes the logical equivalence demonstrating
that \mintinline{coq}{p} is a modal fixed point for \mintinline{coq}{f}.
\begin{minted}{coq}
  Definition fixed_point A G: Prop :=
    forall f, exists p,
    (A; G |-- [! (p <-> f ([]p)) !]).
\end{minted}

>>>>

In this section we describe a case study of the defined modal logic library:
a proof L\"ob theorem in an extension of system K4 with modal fixed points.
The importance of L\"ob's theorem is due to its use in provability
logic~\cite{boolos_1994}, used to abstract details from encodings in G\"odel's
incompleteness results.
%
The purpose of this exercise is to formally prove that the existence of fixed
points (defined below) in a modal logic system implies the GL axiom: for any formula $\Psi$, it
holds that $\Box(\Box\Psi \to \Psi) \to \Box\Psi$. Among the systems defined in
this paper, System GL possesses the fixed point property, though this proof is
non-trivial \cite{fixpoint}; but, in this case, our conclusion becomes rather
uninteresting: the GL axiom is already an axiom in System GL, which trivializes
the result. Still, we justify this case study as other systems that do not have
GL as an axiom but do possess the fixed point property (see \cite{fixpoint}) are
definable, and the main purpose for this code is to formalize a piece of modal
logic proof within Coq by using our framework.

Proceeding,
in order to state L\"ob's theorem we need a way to encode modal
fixed points. Formally, let $\varphi(X)$ be a formula with only one propositional variable
$X$. The modal fixed point of $\varphi$ is then a formula $\Psi$ such that
$\Psi \leftrightarrow \varphi(\Box \Psi)$.
We represent the existence of modal fixed points for all one-variable formulas by a definition \mintinline{coq}{fixed_point},
which is parameterized by a system of modal logic, \mintinline{coq}{A}, and a
theory, \mintinline{coq}{G}, and encodes the logical equivalence demonstrating
that \mintinline{coq}{p} is a modal fixed point for \mintinline{coq}{f}.
\begin{minted}{coq}
  Definition fixed_point A G: Prop :=
    forall f, exists p,
    (A; G |-- [! (p <-> f ([]p)) !]).
\end{minted}
